---
output:
  pdf_document: default
  html_document: default
---
```{r}
# import all the libraries used
library(BSDA)
library(ggplot2)
library(lattice)
library(gridExtra)
library(data.table) # to open large file
library(dplyr)
library(tidyverse)
library(stringi)
library(caret)
library(randomForest)
```

```{r}
## open "black.csv" file using fread
df <- fread("blackfriday.csv")
head(df,5)
```

```{r}
## view five first datas from the dataset
tail(df,5)
```

```{r}
## display the structure the data, such as type, length, dimension, and contents of each column
str(df)
```

```{r}
## display the summary and descriptive statistic of each column
summary(df)
```

```{r}
## view all variables in dataset which are the columns name
colnames(df)
```

```{r}
## check if there are any duplicated values in dataset
sum <- sum(duplicated(df))
sum
## output = 0 --> there are no duplicated values in this dataset
```
```{r}
## convert all the columns name to lowercase for easy use
colnames(df) <- tolower(colnames(df))
colnames(df)
```
```{r}
## columns name has converted to lowercase
head(df)
```

```{r}
## check unique values in gender
unique(df$gender)
```

```{r}
## check unique values in age
unique(df$age)
```

```{r}
## check unique values in occupation
unique(df$occupation)
```

```{r}
## check unique values in city_category
unique(df$city_category)
```

```{r}
## check unique values in stay_in_current_city_years
unique(df$stay_in_current_city_years)
```

```{r}
## replace "4+" with "4" and
df$stay_in_current_city_years[df$stay_in_current_city_years == "4+"] <- "4"

unique(df$stay_in_current_city_years)
```

```{r}
## check unique values in marital_status
unique(df$marital_status)
```

```{r}
## check unique values in product_category_1
unique(df$product_category_1)
```

```{r}
## check unique values in product_category_2
unique(df$product_category_2)
```

```{r}
## check unique values in product_category_3
unique(df$product_category_3)
```

```{r}
## count missing values in product_category_1, product_category_2, and product_category_3
null_count <- colSums(is.na(df[, c("product_category_1", "product_category_2", "product_category_3")]))

print(null_count)
```

```{r}
df2 <- df
```

```{r}
# Replace NaN values with -1 in original dataframe
df$product_category_2 [is.na(df$product_category_2 )] <- -1

# Replace NaN values with mode in original dataframe
df$product_category_3  [is.na(df$product_category_3  )] <- -1

head(df)
```

```{r}
## there are no NaN values
null_count <- colSums(is.na(df[, c("product_category_1", "product_category_2", "product_category_3")]))

print(null_count)
```

```{r}
# get unique values in the 'purchase' column of the 'df' dataframe
unique_vals <- unique(df$purchase)
unique_vals

# get the count of null values in the 'purchase' column of the 'df' dataframe
null_count <- sum(is.na(df$purchase))
null_count
```

```{r}
# Set plot dimensions and adjust margins
options(repr.plot.width=15, repr.plot.height=5)
theme_set(theme_bw(base_size=14))
theme_update(plot.margin=unit(c(1,1,1,1), "cm"))

# Create plot objects
plot1 <- ggplot(df, aes(x=purchase)) + 
  geom_density(fill="blue", alpha=0.3) + 
  xlab("Purchase") +
  ylab("Density") +
  ggtitle("Distribution of Purchase Amounts")

plot2 <- ggplot(df, aes(x=1, y=purchase)) + 
  geom_boxplot(fill="blue", alpha=0.3) +
  xlab("") +
  ylab("Purchase") +
  ggtitle("Boxplot of Purchase Amounts")

# Combine plots into a single figure7
grid.arrange(plot1, plot2, ncol=2)
```
We observed that most of the purchase value is between 5000 to 10000.

```{r}
# group data by 'gender' and count the number of occurrences
gender_counts <- df %>% 
  group_by(gender) %>% 
  summarise(count = n())

# create pie chart to view gender distribution
ggplot(gender_counts, aes(x="", y=count, fill=gender)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  labs(fill="Gender", x=NULL, y=NULL, title="Gender Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust=0.5)) +
  geom_text(aes(label = paste0(round(count/sum(count)*100),"%")),
            position = position_stack(vjust = 0.5)) +
  guides(fill=guide_legend(title="Gender")) +
  theme(legend.position="right") 
```
Customers who did the purchase are 75% male.

```{r}
ma <- df[df$gender == 'M',]
fe <- df[df$gender == 'F',]

p1 <- ggplot(ma, aes(x = purchase)) + 
  geom_density(fill = "blue", alpha = 0.3) +
  ggtitle("Male")

p2 <- ggplot(fe, aes(x = purchase)) +
  geom_density(fill = "pink", alpha = 0.3) +
  ggtitle("Female")

grid.arrange(p1, p2, ncol = 2)
```
Purchase distribution of male and female.

```{r}
p <- df %>% 
  group_by(gender) %>%
  summarize(purchase = sum(purchase)) %>%
  ungroup()
ggplot(p, aes(x = gender, y = purchase)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "Gender", y = "Total Purchase") +
  ggtitle("Total Purchase by Gender")
```

Male customers purchased more.
```{r}
# Create frequency table and sort values in descending order
p_id <- df %>% 
  count(product_id) %>% 
  arrange(desc(n)) %>% 
  head(40) %>% 
  arrange(desc(n))

# Create a bar plot
ggplot(data = p_id, aes(x = reorder(product_id, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() + labs(x = "Product ID", y = "Count") + coord_flip()
```
Top 40 products purchased by customers.

```{r}
c1 <- df %>% 
  group_by(gender, product_category_1) %>% 
  summarise(p_count = n()) %>% 
  ungroup()

ggplot(c1, aes(x = product_category_1, y = p_count, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(x = "Product Category 1", y = "Count", fill = "Gender")
```
The top purchased products from category 1 is category 1, 5, and 8. Most of the customers who bought those products are male.

```{r}
c2 <- df %>% 
  group_by(gender, product_category_2) %>% 
  summarise(p_count = n()) %>% 
  ungroup()

ggplot(c2, aes(x = product_category_2, y = p_count, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(x = "Product Category 2", y = "Count", fill = "Gender")
```
The top purchased products from category 2 is category 8. Most of the customers who bought those products are male.

```{r}
c3 <- df %>% 
  group_by(gender, product_category_3) %>% 
  summarise(p_count = n()) %>% 
  ungroup()

ggplot(c3, aes(x = product_category_3, y = p_count, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(x = "Product Category 3", y = "Count", fill = "Gender")
```
The top purchased products from category 3 is category 16. Most of the customers who bought those products are male.

```{r}
df %>% 
  count(marital_status) %>% 
  ggplot(aes(x = "", y = n, fill = factor(marital_status))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  scale_fill_manual(values = c("steelblue", "orange")) +
  geom_text(aes(label = paste0(round(n/sum(n)*100), "%")), position = position_stack(vjust = 0.5)) +
  labs(fill = "Marital Status") +
  ggtitle("Marital Status")
```

Unmarried customers purchased more than married customers.
```{r}
UM_P <- df %>% 
  filter(marital_status == 0) %>% 
  count(purchase) %>% 
  rename(count = n)

ggplot(UM_P, aes(x = purchase)) +
  geom_density(fill = "steelblue") +
  theme_minimal() +
  labs(x = "Purchase", y = "Density", title = "UnMarried")
```

Purchase distribution of unmarried customers.
```{r}
M_P <- df %>% 
  filter(marital_status == 1) %>% 
  count(purchase) %>% 
  rename(count = n)

ggplot(M_P, aes(x = purchase)) +
  geom_density(fill = "steelblue") +
  theme_minimal() +
  labs(x = "Purchase", y = "Density", title = "Married")
```

Purchase distribution of married customers.
```{r}
M_UM_P <- df %>% 
           group_by(gender, marital_status) %>%
           summarise(purchase = sum(purchase)) %>%
           ungroup()

ggplot(M_UM_P, aes(x = marital_status, y = purchase, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge")
```

Unmarried customers purchased more with most of them are male.
```{r}
UM <- df %>%
        filter(marital_status == 0) %>%
        count(product_id) %>%
        arrange(desc(n)) %>%
        head(10)

ggplot(UM, aes(x = product_id, y = n)) +
  geom_bar(stat = "identity") +
  labs(title = "UnMarried_Customer") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Top 10 purchased product id by unmarried customers.
```{r}
M <- df %>%
       filter(marital_status == 1) %>%
       count(product_id) %>%
       arrange(desc(n)) %>%
       head(10)

ggplot(M, aes(x = product_id, y = n)) +
  geom_bar(stat = "identity") +
  labs(title = "Married_Customer") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Top 10 purchased product id by married customers.
```{r}
c1 <- df %>% 
  group_by(purchase, product_category_1) %>% 
  summarise(p_count = n()) %>% 
  ungroup()

ggplot(c1, aes(x = product_category_1, y = p_count, fill = purchase)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(x = "Product Category 1", y = "Total Purchase", fill = "Purchase")
```

The most purchased product from category 1 is product 5, followed by product 8, and product 19.


```{r}
c3 <- df %>% 
  group_by(purchase, product_category_3) %>% 
  summarise(p_count = n()) %>% 
  ungroup()

ggplot(c3, aes(x = product_category_3, y = p_count, fill = purchase)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(x = "Product Category 3", y = "Total purchase", fill = "Purchase")
```
The most purchased product from category 3 is product 14.

===============
| CORRELATION |
===============

```{r}
str(df)
```

Dependent Variable: Total Purchase (continuous)
Independent Variable: 
- Gender (nominal) - Point Biserial
- Age (continuous) - Spearman
- City Category (nominal) - Spearman
- Stay in Current City years (ordinal) - Pearson / Spearman
- Marital status (nominal) - Point Biserial

==========
| GENDER |
==========
```{r}
# replace "M" with "0" and "F" with "1"
df$gender[df$gender == "M"] <- "0"
df$gender[df$gender == "F"] <- "1"

# Specify the dependent variable and the independent variable of interest
dep_var_purchase <- as.numeric(df$purchase)
indep_var_gender <- as.numeric(df$gender)

# Calculate the Point Biserial correlation
cor_result_gender <- cor(dep_var_purchase, indep_var_gender)

# View the correlation results
cor_result_gender
```

```{r}
dfs <- df$purchase
x <- dfs
# get median
med = median(x)
# subtract median from each value of x and get absolute deviation
abs_dev = abs(x-med)
# get MAD
mad = 1.4826 * median(abs_dev)
# get threshold values for outliers
Tmin = med-(3*mad) 
Tmax = med+(3*mad) 
# find outlier
x[which(x < Tmin | x > Tmax)]
y = x[which(x > Tmin & x < Tmax)]
```

```{r}
# Replace outliers with the mean
mean_value <- mean(df$purchase,na.rm=TRUE)
df$purchase <- ifelse(df$purchase < Tmin | df$purchase > Tmax, mean_value, df$purchase)

dep_var_purchase <- as.numeric(df$purchase)
indep_var_gender <- as.numeric(df$gender)

# Calculate the Point Biserial correlation
cor_result_gender <- cor(dep_var_purchase, indep_var_gender)

# View the correlation results
cor_result_gender
```

=======
| Age |
=======
```{r}
# change the age range using the median so you can find the correlation
df$age[df$age == "0-17"] <- "0"
df$age[df$age == "18-25"] <- "1"
df$age[df$age == "26-35"] <- "2"
df$age[df$age == "36-45"] <- "3"
df$age[df$age == "46-50"] <- "4"
df$age[df$age == "51-55"] <- "5"
df$age[df$age == "55+"] <- "6"

# Specify the dependent variable and the independent variable of interest
dep_var_purchase <- as.numeric(df$purchase)
indep_var_age <- as.numeric(df$age)

# Calculate correlations using Pearson's correlation coefficient
cor_results_age <- cor(dep_var_purchase, indep_var_age, method = "spearman")

# Print the correlation coefficient
cor_results_age
```

=================
| City Category |
=================
```{r}
# Perform ANOVA
anova_results <- aov(df$purchase ~ df$city_category)

# Print the ANOVA table
summary(anova_results)

# Perform Kruskal-Wallis test
kruskal_results <- kruskal.test(df$purchase ~ df$city_category)

# Print the Kruskal-Wallis test results
kruskal_results
```

```{r}
# Calculate Spearman's correlation coefficient
cor_results_city <- cor(as.numeric(as.factor(df$city_category)), df$purchase, method = "spearman")

# Print the correlation coefficient
cor_results_city
```

==============================
| Stay in Current City years |
==============================
```{r}
# Calculate Spearman's rank correlation
cor_results_stay <- cor(as.numeric(df$stay_in_current_city_years), df$purchase, method = "spearman")

# Print the correlation coefficient
cor_results_stay
```

==================
| Marital Status |
==================
```{r}
# Specify the dependent variable and the independent variable of interest
dep_var_purchase <- as.numeric(df$purchase)
indep_var_marital <- as.numeric(df$marital_status)

# Calculate the Point Biserial correlation
cor_result_marital <- cor(dep_var_purchase, indep_var_marital)

# View the correlation results
cor_result_marital
```

=============================
| Statistical Test (T-Test) |
=============================
```{r}
result_1 <- t.test(df$product_category_1, df$product_category_2)
result_1
```

=============================
| Statistical Test (Z-Test) |
=============================
```{r}
sd_1 <- sd(df$product_category_1)
sd_2 <- sd(df$product_category_2)
result_2 <- z.test(df$product_category_1, df$product_category_2, 
                   sigma.x = sd_1, sigma.y = sd_2)

result_2
```

```{r}
unique(df$occupation)
df$user_id <- as.numeric(factor(df$user_id, levels = unique(df$user_id))) - 1
df$product_id <- as.numeric(factor(df$product_id, levels = unique(df$product_id))) - 1
df$city_category <- as.numeric(factor(df$city_category, levels = unique(df$city_category))) - 1
df$product_category_1 <- as.numeric(factor(df$product_category_1, levels = unique(df$product_category_1))) - 1
df$product_category_2 <- as.numeric(factor(df$product_category_2, levels = unique(df$product_category_2))) - 1
df$product_category_3 <- as.numeric(factor(df$product_category_3, levels = unique(df$product_category_3))) - 1

df
```

```{r}
str(df)
```

```{r}
df$user_id <- as.integer(df$user_id)
df$product_id <- as.integer(df$product_id)
df$gender <- as.integer(df$gender)
df$age <- as.integer(df$age)
df$gender <- as.integer(df$gender)
df$city_category <- as.integer(df$city_category)

df$stay_in_current_city_years <- as.integer(df$stay_in_current_city_years)
df$occupation <- as.integer(df$occupation)
df$marital_status <- as.integer(df$marital_status)
df$product_category_1 <- as.integer(df$product_category_1)
df$product_category_2 <- as.integer(df$product_category_2)
df$product_category_3 <- as.integer(df$product_category_3)
df$purchase <- as.integer(df$purchase)

str(df)
```

=====================
| Linear Regression |
=====================
```{r}
# Membagi dataset menjadi set pelatihan dan set pengujian
set.seed(123)  # Untuk reproduksi hasil yang sama
training_index <- sample(1:nrow(df), 0.8*nrow(df))  # Mengambil 80% data sebagai set pelatihan
training_set <- df[training_index, ]
testing_set <- df[-training_index, ]

# Membangun model regresi linier
model <- lm(purchase ~ gender + age + occupation + city_category + stay_in_current_city_years + product_category_1 + product_category_2 + product_category_3 + marital_status , data = training_set)  # Menggunakan semua variabel independen untuk memprediksi variabel "purchase"

summary(model)

# Melakukan prediksi pada set pengujian
predictions <- predict(model, newdata = testing_set)

# Evaluasi performa model
mse <- mean((testing_set$purchase - predictions)^2)  # Mean Squared Error
mae <- mean(abs(testing_set$purchase - predictions))  # Mean Absolute Error
rsquared <- 1 - sum((testing_set$purchase - predictions)^2) / sum((testing_set$purchase - mean(testing_set$purchase))^2)  # R-squared

# Menampilkan hasil evaluasi
print(paste("Mean Squared Error:", mse))
print(paste("Mean Absolute Error:", mae))
print(paste("R-squared:", rsquared))
```
```{r}
df
```

=================
| Random Forest |
=================
```{r}
library(randomForest)

# Menghapus kolom yang tidak diperlukan
df3 <- subset(df, select = -c(product_id, user_id))

# Menentukan variabel prediktor dan target
X <- df3[, c("gender", "age", "occupation", "city_category", "marital_status", "stay_in_current_city_years", "product_category_1", "product_category_2", "product_category_3")]
Y <- df3$purchase

# Splitting the dataset
set.seed(100)  # Set random seed for reproducibility
train_indices <- sample(1:nrow(df3), 0.1 * nrow(df3))
X_train <- X[train_indices, ]
y_train <- Y[train_indices]
X_test <- X[-train_indices, ]
y_test <- Y[-train_indices]
# Model Training
rfr <- randomForest(X_train, y_train)

# Training Accuracy
print(rfr)

# Testing Accuracy
y_pred <- predict(rfr, X_test)
rscore <- cor(y_test, y_pred)^2
print(paste("R-squared:", rscore))

# Predictions and Error Calculation
predictions <- data.frame(Prediction = y_pred, Actual = y_test, Error = y_test - y_pred)
print(predictions)
```
